---
title: "analysis"
author: "Parker Brotman"
date: '2023-01-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(sf)
```


# Load data

## Starbucks data
```{r}
tbl_starbucks <- read_csv("../data/starbucks/directory.csv") %>%
  filter(Country=="US")
```

### Find county for starbucks
Convert starbucks long/lat to sf points
```{r}
pts_starbucks <- st_as_sf(tbl_starbucks, coords = c("Longitude", "Latitude"), crs = 4326) #sets/reads format to WGS 84
```

Load county polygons
```{r}
shape_county <- st_read("../data/census/Geographic Shape Files/cb_2018_us_county_500k/cb_2018_us_county_500k.shp") %>%
  st_transform(crs = 4326)
```
Left join Starbucks and county
```{r}
tbl_starbucks_w_county <- st_join(pts_starbucks, shape_county)
```

Check: View Starbucks in MD
```{r}
ggplot(shape_county %>% filter(STATEFP=="24"))+
  geom_sf(aes())+
  geom_sf(data=tbl_starbucks_w_county %>% filter(STATEFP=="24"))
```


## Census data

### Race
```{r}
tbl_race_raw <- read_csv("../data/census/Race/County Level/DECENNIALPL2020.P1-Data.csv", skip = 1) 

tbl_race <- tbl_race_raw%>% 
  select(-matches("^Annotation")) %>%
  rename_with(.fn = ~str_extract(.x, "(?<=\\!\\!)[^\\!\\:]*(?=\\:?$)"), .cols = starts_with("!"))
```

### Region/Division
```{r}
tbl_regions <- read_csv("../data/census/Regions/regions.csv")
```


# Analysis

## Data Joining

Tally # Starbucks per county
```{r}
tbl_starbucks_per_county <- tbl_starbucks_w_county %>% 
  group_by(AFFGEOID, NAME, `State/Province`) %>% 
  summarise(n = n()) %>% 
  arrange(desc(n))
tbl_starbucks_per_county
```

Join with race data
```{r}
tbl_sb_race <- full_join(tbl_starbucks_per_county, tbl_race, by=c("AFFGEOID"="Geography")) %>%
  filter(!is.na(Total)) %>% 
  mutate(n = replace_na(n, 0)) %>% 
  separate(`Geographic Area Name`, c("County", "State"), ", ") %>% 
  rename("State Code" = "State/Province") %>% 
  select(-"...145") %>% 
  filter(State!="Puerto Rico")
tbl_sb_race
```

Join with regions table
```{r}
tbl_sb_race_region <- tbl_sb_race %>% 
  select(-`State Code`) %>% 
  left_join(tbl_regions, by=c("State"))
```


## Models

### Total pop
```{r}
ggplot(tbl_sb_race, aes(x=Total, y=n))+
  geom_point()+
  stat_smooth(method = "lm", col = "red")
```


```{r}
lm_pop <- lm(n ~ Total, data=tbl_sb_race)
# lm_pop <- lm(n ~ Total, data=tbl_sb_race %>% filter(n<400))
summary(lm_pop)
plot(lm_pop)
# Distribution is heavy-tailed
```

#### Analyze residuals

Residual by county
```{r}
tbl_res.lm_pop <- tbl_sb_race %>% 
  st_set_geometry(NULL) %>% 
  add_column(pred = predict(lm_pop),
             res = resid(lm_pop)) %>% 
  full_join(shape_county, by = "AFFGEOID") %>% 
  select(AFFGEOID, County, State, n, Total, pred, res, geometry)
```

Residuals geographic view
```{r}
tbl_res.lm_pop %>% 
  # filter(State %in% c("District of Columbia", "Maryland", "Virginia")) %>% 
  filter(State=="Maryland") %>% 
  st_set_geometry("geometry") %>% 
  ggplot()+
  geom_sf(aes(fill=res))
```
Largest residuals
```{r}
tbl_res.lm_pop %>% 
  ungroup() %>% 
  select(-AFFGEOID) %>% 
  arrange(desc(res))
```

Smallest residuals
```{r}
tbl_res.lm_pop %>% 
  ungroup() %>% 
  select(-AFFGEOID) %>% 
  arrange(res)
```

### Race
```{r}
lm_race <- lm(n ~ Total + `White alone` + `Black or African American alone` + `Asian alone`, data=tbl_sb_race)
summary(lm_race)
plot(lm_race)
```

### What about predicting Starbucks per capita (doesn't work as well)
```{r}
tbl_sb_race_pct <- tbl_sb_race %>% 
  # rename_all(~str_replace_all(., " ", "_")) %>% 
  select(-Total, Total) %>% # need to put total at end, otherwise next columns are divided by 1
  mutate_if(~is.numeric(.), ~./Total)
```


```{r}
lm_pop_pct <- lm(n ~ `White alone` + `Black or African American alone` + `Asian alone`, data=tbl_sb_race_pct)
summary(lm_pop_pct)
plot(lm_pop_pct)
```

#### Likewise, log doesn't work
Doesn't work in any combination of log.
```{r, eval=FALSE}
lm_pop_log <- lm(log1p(n) ~ log1p(Total), data=tbl_sb_race)
summary(lm_pop_log)
plot(lm_pop_log)
```

### Regions/Divisions as predictor
```{r}
ggplot(tbl_sb_race_region, aes(x=Total, y=n))+
  facet_wrap(~Division)+
  geom_point()+
  stat_smooth(method = "lm", col = "red")
```

```{r}
lm_pop_region <- lm(n ~ Total*Region, data=tbl_sb_race_region) # Midwest in reference
summary(lm_pop_region)
plot(lm_pop_region)
```

```{r}
lm_pop_division <- lm(n ~ Total*Division, data=tbl_sb_race_region) # East North Central is reference
summary(lm_pop_division)
plot(lm_pop_division)
```

Residual by county
```{r}
tbl_res.lm_pop_division <- tbl_sb_race_region %>% 
  st_set_geometry(NULL) %>% 
  add_column(pred = predict(lm_pop_division),
             res = resid(lm_pop_division)) %>% 
  full_join(shape_county, by = "AFFGEOID") %>% 
  select(AFFGEOID, County, State, Division, n, Total, pred, res, geometry)
```

```{r}
tbl_res.lm_pop_division %>% 
  ungroup() %>% 
  select(-AFFGEOID) %>% 
  arrange(desc(res))
```

```{r}
tbl_res.lm_pop_division %>% 
  ungroup() %>% 
  select(-AFFGEOID) %>% 
  arrange(res)
```