---
title: "Analysis 3.0"
author: "Parker Brotman"
date: '2023-01-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(sf)
```

# Load data

## Starbucks data
```{r}
tbl_starbucks <- read_csv("../data/starbucks/directory.csv") %>%
  filter(Country=="US")
```

## Census data

While I considered a more generalized read design that reads every file in a specified folder, I opted instead to read each table individually, since ultimately, I'll need to subset the # of variables anyway.

### Define functions
```{r}
fn_load_decennial <- function(fname){
  str_c("../data/census/Decennial_2020/County Level/", fname) %>% 
    read_csv(skip = 1) %>% 
    select(-matches("^Annotation")) %>% 
    rename_with(.fn = ~str_extract(.x, "(?<=\\!\\!)[^\\!\\:]*(?=\\:?$)"), .cols = starts_with("!")) %>% # Preceded by !!, some number of non-!: characters, followed by end line or : + end 
    return()
}
```

```{r}
rename_prefix <- function(tbl, old_prefix, new_prefix){
  tbl %>% 
    rename_with(.fn = ~str_c(new_prefix, str_extract(.x, "(?<=\\!\\!)[^\\!\\:]*(?=$)")), .cols = starts_with(old_prefix)) %>% 
    return()
}
```


### Race & Hispanic/Latino
```{r}
tbl_race <- fn_load_decennial("Race (DECENNIALPL2020.P1-Data).csv") %>% 
  select(Geography:`Population of two or more races`)

tbl_hisp_latino <- fn_load_decennial("Hispanic-Latino (DECENNIALPL2020.P2-Data).csv") %>% 
  select(Geography:`Not Hispanic or Latino`)

list_tbl_decennial <- list(race = tbl_race, hisp_latino = tbl_hisp_latino)
```


### Age & Sex
```{r}
tbl_age_sex <- read_csv("../data/census/ACS_2021/County Level/Age-Sex (ACSST5Y2021.S0101-Data).csv", skip = 1) %>% 
  select(-matches("^Annotation|^Margin of Error|^Estimate\\!\\!Percent")) %>%
  rename_prefix("Estimate!!Total", "") %>% 
  rename_prefix("Estimate!!Male", "Male & ") %>%
  rename_prefix("Estimate!!Female", "Female & ") %>% 
  select(Geography:`85 years and over`, `Male & Total population`, `Female & Total population`) %>% 
  rename(Male=`Male & Total population`, 
         Female=`Female & Total population`)
```

### Education
```{r}
tbl_education <- read_csv("../data/census/ACS_2021/County Level/Education (ACSST5Y2021.S1501-Data).csv", skip = 1) %>% 
  select(Geography, `Geographic Area Name`, matches("^Estimate!!Total!!AGE BY EDUCATIONAL ATTAINMENT!!Population 25 years and over")) %>% 
  rename_prefix("Estimate!!Total", "") %>% 
  select(Geography:`Graduate or professional degree`, -`Population 25 years and over`)
```

### Households
Note: relevant data only available in % of total households (create intereaction term with population to make it scale).
```{r}
tbl_household <- read_csv("../data/census/ACS_2021/County Level/Household (ACSST5Y2021.S1101-Data).csv", skip = 1) %>% 
  select(Geography, `Geographic Area Name`, matches("^Estimate!!Total!!Total Households!!UNITS IN STRUCTURE|^Estimate!!Total!!Total Households!!HOUSING TENURE")) %>% 
  rename_prefix("Estimate!!Total", "% in ")
```

### Income
Note: relevant data only available in % of total households (create intereaction term with population to make it scale).
```{r}
tbl_income <- read_csv("../data/census/ACS_2021/County Level/Income (ACSST5Y2021.S1901-Data).csv", skip = 1) %>% 
  select(Geography, `Geographic Area Name`, matches("^Estimate!!Households!!Total"), -`Estimate!!Households!!Total`) %>% 
  rename_prefix("Estimate", "% ")
```



# Combine data